---
title: "RF to precit the manner of do exercise"
author: "Sebastian Barreto"
date: "September of 2020"
output: html_document
---

## The Data

First the data sets are uploaded to the R environment. It is necessary to consider the empty data at the moment of uploading the databases, so the bla instruction **na.strings=c("","NA","#DIV/0!")** allows to characterize empty fields, with calculation error OR NAs as NAs

```{r 1, include=T}
data.Training <-read.csv("D:/Coursera/Practical Machine Learning/Proyecto/pml-training.csv", na.strings=c("","NA","#DIV/0!"))
data.Test <-read.csv("D:/Coursera/Practical Machine Learning/Proyecto/pml-testing.csv", na.strings=c("","NA","#DIV/0!"))
```

The following is to eliminate all those columns that present data with NAs, and select only those variables that can provide useful information to the model

```{r 2, include=T}
library(caret)
data.Training<- data.Training[ , colSums(is.na(data.Training)) == 0]
data.Training <- data.Training[, c(8:60)]
data.Test<- data.Test[ , colSums(is.na(data.Test)) == 0]
```

## Training, validation and set data

The data set **data.Training** will be separated into two subsets: the model training set and the model validation set:

```{r 3, include=T}
data.Partition <- createDataPartition(y=data.Training$classe,
                                    p=0.7, list=F)
subset.Train <- data.Training[data.Partition,]
subset.Validation <- data.Training[-data.Partition,]
```

A third set of data (**Test**) will serve to test the model, predicting the type of exercise in 20 observations of interest:

```{r 4, include=T}
subset.Test <- data.Test
subset.Test <- subset.Test[, c(8:59)]
```

## Random Forest model

For training the *RF**** model:

```{r 5, include=T}
mod.RF <- train(classe ~ ., data=subset.Train,
                method="rf")
```

See model results:

```{r 6, include=T}
mod.RF
```

Los resultados del modelo indican que la mayor precisión (**0.9883115**) se alcanza al utilizar un valor de **2** para el parámetro **mtry**

Podemos ver la importancia de las variables dentro del modelo con:

```{r d, include=T}
gbmImp <- varImp(mod.RF, scale = FALSE)
gbmImp
```

The 5 most important variables are: 1) roll_belt , 2) yaw_belt, 3) magnet_dumbbell_z, 4) pitch_forearm and 5) magnet_dumbbell_y   

## Validation and prediction on test

```{r 7, include=T}
prediction.Validation = predict(mod.RF, newdata = subset.Validation)
confusionMatrix(prediction.Validation, subset.Validation$classe)
```

The result of the validation process with the validation set gives an accuracy of 0.9952, which allows us to indicate that it is a model of very high reliability

To perform the class prediction on the Test data set we use

```{r 8, include=T}
prediction.Test = predict(mod.RF, newdata = subset.Test)
```

Te result of the predictiopn is:

```{r sd, include=T}
prediction.Test
```